---
title: "KBB: Individualized Report"
author: ""
date: ""
output: html_document
---

```{r setup, echo = F}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)
knitr::opts_chunk$set(warning = FALSE)

```

```{r loading in the packages, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(readxl)
library(kableExtra)
```

```{r pick the ID number}
ID_num = 41 #41 #62 

```


```{r loading in all required data}
# Set working directory to where your GitHub repository is
DataLocation = "~/Documents/Github/KBB/"

# Set working directory to where the final data is
setwd('C:/Users/lledesma.TIMES/Documents/KBB/Data/FINAL_DS/Behavioral')

# Load in the full data
full_data <- read_excel("full_data.xlsx")

# drop any NAs in the dataset for Child_ID
full_data <- drop_na(full_data, Child_ID)
```

```{r load in dd status data}
# Set the location for where the ID tracker is saved and the other where the CFM data is saved
DataLocation2 = "C:/Users/lledesma.TIMES/Documents/KBB/Data/FINAL_DS/Screener/Matched_Siblings/"
DataLocation3 = "C:/Users/lledesma.TIMES/Documents/KBB/Data/FINAL_DS/Screener/"

# Load tracker ID infor
setwd(DataLocation2)
trackerID <- read_excel("Final_ID_Tracker.xlsx")

# Load the CFM data
setwd(DataLocation3)
CFM <- read_excel("All Children.xlsx")

# Some minor data cleaning
trackerID2 <- select(trackerID, full_child_ID = Child_ID, Child_ID = ID)
CFM2 <- select(CFM, full_child_ID = Child_ID, KBB_CFM_DD_type)

# Merge the datasets by child_ID
CFM_ID <- trackerID2 %>%
  full_join(CFM2, by = "full_child_ID") %>%
  select(-full_child_ID)

# Introdude this info into full data
full_data <- full_data %>%
  full_join(CFM_ID, "Child_ID")
  
# drop NA's in Child_ID
full_data <- drop_na(full_data, Child_ID)
```

```{r create groups by age}
# Arbitrary groups
# 2-4;
# 5-8;
# 9-12; 
# 13-15;
# 16-19 

full_data <- full_data %>%
  mutate(age_group = case_when(
     Age >= 2 & Age < 5 ~ "g2_4",
     Age >= 5 & Age < 9 ~ "g5_8",
     Age >= 9 & Age < 13 ~ "g9_12",
     Age >= 13 & Age < 16 ~ "g13_15",
     Age >= 16 & Age < 19 ~ "g16_19",
     TRUE ~ "ERROR"
  ))

# Drop any ERROR (kids have not have data collected yet)
full_data <- filter(full_data, age_group != "ERROR")
```


```{r obtain zscores for each task for each group}
full_data_z <- full_data %>%
  group_by(age_group) %>%
  mutate(KBB_CFM_DD_type,
         Epilepsy,
         PSC_Performance_z = c(scale(PSC_Performance)),
         CBC3_6_emotionally_reactive_z = c(scale(CBC3_6_emotionally_reactive)),
         CBC3_6_anxious_depressed_z = c(scale(CBC3_6_anxious_depressed)),
         CBC3_6_somatic_complaints_z = c(scale(CBC3_6_somatic_complaints)),
         CBC3_6_withdrawn_z = c(scale(CBC3_6_withdrawn)),
         CBC3_6_sleep_problems_z = c(scale(CBC3_6_sleep_problems)),
         CBC3_6_attention_problems_z = c(scale(CBC3_6_attention_problems)),
         CBC3_6_aggressive_behaviors_z = c(scale(CBC3_6_aggressive_behaviors)),
         CBC3_6_other_problems_z = c(scale(CBC3_6_other_problems)),
         CBC3_6_Internalizing_z = c(scale(CBC3_6_Internalizing)),
         CBC3_6_Externalizing_z = c(scale(CBC3_6_Externalizing)),
         CBC3_6_Total_Prob_z = c(scale(CBC3_6_Total_Prob)),
         CBC3_6_depressive_problems_z = c(scale(CBC3_6_depressive_problems)),
         CBC3_6_anxiety_problems_z = c(scale(CBC3_6_anxiety_problems)),
         CBC3_6_autism_spectrum_problems_z = c(scale(CBC3_6_autism_spectrum_problems)),
         CBC3_6_attention_deficit_hyperactivity_problems_z = c(scale(CBC3_6_attention_deficit_hyperactivity_problems)),
         CBC3_6_opositional_defiant_problems_z = c(scale(CBC3_6_opositional_defiant_problems)),
         CBC6_18_anxious_depressed_z = c(scale(CBC6_18_anxious_depressed)),
         CBC6_18_withdrawn_depressed_z = c(scale(CBC6_18_withdrawn_depressed)),
         CBC6_18_somatic_complaints_z = c(scale(CBC6_18_somatic_complaints)),
         CBC6_18_social_problems_z = c(scale(CBC6_18_social_problems)),
         CBC6_18_thought_problems_z = c(scale(CBC6_18_thought_problems)),
         CBC6_18_attention_problems_z = c(scale(CBC6_18_attention_problems)),
         CBC6_18_ruleBreaking_behavior_z = c(scale(CBC6_18_ruleBreaking_behavior)),
         CBC6_18_aggressive_behavior_z = c(scale(CBC6_18_aggressive_behavior)),
         CBC6_18_other_problems_z = c(scale(CBC6_18_other_problems)),
         CBC6_18_Internalizing_z = c(scale(CBC6_18_Internalizing)),
         CBC6_18_Externalizing_z = c(scale(CBC6_18_Externalizing)),
         CBC6_18_Total_Prob_z = c(scale(CBC6_18_Total_Prob)),
         RV_Performance_z = c(scale(RV_Performance)),
         PR_Performance_z = c(scale(PR_Performance)),
         TR_Performance_z = c(scale(TR_Performance)),
         LetDig_Performance_z = c(scale(LetDig_Performance))) %>%
  ungroup()

```


```{r filter data by participant ID}
data <- full_data_z %>%
  filter(Child_ID == ID_num)

cat("The following report is for the child with the ID: ", ID_num, sep = "")
```
### Demographics

Below is the age and sex of the participant.

```{r print out the demographics of the current child}
select(data, Age, Sex) %>%
  kbl() %>%
  kable_minimal(full_width = F)


```


```{r create a function to report zscore results}
# We decided on the following categories
# Bottom 5% → "Lowest individuals."
# 5% to 20% → "Below average individuals."
# 20% to 80% → "Average individuals."
# 80% to 95% → "Above average individuals."
# Top 5% → "Highest individuals."

zscore_fun <- function(x) {
  if (is.na(x)) {
    return(NA_character_)  # Ensures NA is returned as a character
  }
  
  if (x >= 1.645) {
    return("Top 5% - Highest")
  } else if (x >= 0.84) {
    return("80-95% - Above average")
  } else if (x >= -0.84) {
    return("20-80% - Average")
  } else if (x >= -1.645) {
    return("5-20% - Below average")
  } else {
    return("Bottom 5% - Lowest")
  }
}


```


### CFM DD and Epilepsy Status

Below are the domains from the child functioning module (CFM) that the child were reported to have difficulties in. The table shows all domains that "at least some difficulty" was reported for. Additionally, epilepsy status is also available based on information from the **Birbeck Epilepsy Screener** at the bottom of the table. 

```{r print out CFM DD status by severity and epsilepsy status}
# Obtain variables of interest
KBB_CFM <- data %>%
  select(KBB_CFM_DD_type, Epilepsy)

# Report the CFM domains that at least some difficulty was reported
KBB_CFM2 <- data.frame(Domains = do.call(c, str_split(KBB_CFM$KBB_CFM_DD_type, "; "))) %>%
  mutate(`Difficulty Reported` = "Yes")
  
# Add Epilepsy info into the dataframe
KBB_CFM3 <- rbind(KBB_CFM2, tibble(Domains = "Epilepsy",
                           `Difficulty Reported` = KBB_CFM$Epilepsy))

# For nonDD, if Domains is = NA replace reported difficulty to No
if(is.na(KBB_CFM3$Domains[1])) {
  KBB_CFM3$`Difficulty Reported`[1] <- "No Difficulties Reported"
  
}

# Print out the table
KBB_CFM3 %>%
  kbl() %>%
  kable_styling(full_width = F)
```

### Physical Data

**Eye sight**: The Snelle chart system was use to measure seeing ability. The score is represented as a fraction in which the **numerator** represents the distance in feet the child needs to see a letter or symbol while the **denominator** represents the distance in feet the average child needs to see that same letter or symbol. The numerator is typically fixed to the value 20, thus what decides visual acuity is the **denominator**. Therefore, if the denominator is smaller than 20, then the child tends to have better eye-sight because they can see something well from 20 feet that will require the average child a shorter distance to view it. However, if the denominator is larger, then that indicates the child has poorer vision since it takes them 20 feet to view a symbol that the average child can see from more than 20 feet away.

Summary Examples:

- 20/16 = better than normal vision
- 20/20 = normal vision
- 20/25 = slightly worse than normal, but still functional
- 20/40 = acceptable for daily activities, but not ideal
- 20/100 = significant vision impairment
- 20/200 and worse = legally blind or very poor vision

```{r physical data eyes}
Eyes <- data %>%
  select(Left.eye:Both.Eyes)

# Print the tables
Eyes %>%
  kbl() %>%
  kable_styling(full_width = F)

```

**Hearing**: A hearing test was performed to identify the decibels (dB) needed for the child to perceive (hear) three different frequencies (1000Hz, 2000Hz, and 4000 Hz) in both ears. Ideally the smaller the decibal that it takes to perceive the sound, the better the hearing. 

Summary Examples:

- -10 to -1 - better than normal hearing:
- 0 to 20 dB - normal hearing:
- 21 to 40 dB - mild hearing loss: may have trouble hearing soft sounds like whispers
- 41 to 70 dB - moderate hearing loss: may have difficulty hearing normal conversations
- 71 to 90 dB - severe hearing loss: may only hear loud speech or sound
- 91 dB+ - profound hearing loss: may not be able to hear everyday sounds without hearing aids
  

```{r physical data hearing}
Ears_L <- data %>%
  select(`1000Hz_dB` = `1000Hz_Left_dB`,
         `2000Hz_dB` = `2000Hz_Left_dB`,
         `4000Hz_dB` = `4000Hz_Left_dB`,) %>%
  mutate(Ear = "Left")

Ears_R <- data %>%
  select(`1000Hz_dB` = `1000Hz_Right_dB`,
         `2000Hz_dB` = `2000Hz_Right_dB`,
         `4000Hz_dB` = `4000Hz_Right_dB`) %>%
  mutate(Ear = "Right")

Ears <- rbind(Ears_L, Ears_R) %>% select(Ear, everything())


Ears %>%
  kbl(caption = "") %>%
  kable_styling(full_width = F)
```

### Assessment Scoring

For the remaining assessments a z-score was calculated from each raw score within specific age groups specified below. Thus, only the scores of children within an age group were compared to each other- this controls for performance improving by an increase in age. 

```{r group sample size}
# Create a dataframe for the groups created
group_n <- data.frame(table(full_data$age_group))

# Clean the data a bit
group_n2 <- rename(group_n, Group = Var1, n = Freq)
group_n2$Group <- gsub("g", "",group_n2$Group)
group_n2$order <- as.numeric(sapply(str_split(group_n2$Group, "_"), function(x) x[1]))
group_n2 <- arrange(group_n2, order)
group_n3 <- select(group_n2, -order)


# Identify how many children with ID have full data
older <- full_data %>%
  filter( Age > 5.9) %>%
  select(Age, PSC_Performance, CBC6_18_anxious_depressed:LetDig_Performance, age_group) %>%
  filter(complete.cases(.))

older2 <- select(older, age_group)

younger <- full_data %>%
  filter( Age < 6) %>%
  select(Age, PSC_Performance, CBC3_6_emotionally_reactive:CBC3_6_opositional_defiant_problems , RV_Performance:LetDig_Performance, age_group) %>%
  filter(complete.cases(.))

younger2 <- select(younger, age_group)

# Combine these into one dataset
available_data <- rbind(older2, younger2)

available_data_df <- data.frame(table(available_data$age_group))

# Cheese to get the order I want 
group_n3 %>%
  mutate(Group = gsub("_","-", Group)) %>%
  rename(`Age Group` = Group) %>%
  kbl() %>%
  kable_minimal(full_width = F)

```



The z-scores represent how far the child's score is away from the mean (the average of the sample), with the value -1 being one standard deviation below the mean, a value of 0 indicating a score at the mean, and a value of 1 indicating a score one standard deviation above the mean. These z-scores can be converted into percentiles to create categories indicating how the child performed compared to their peers in the same age group. The categories chosen are the following:

```{r printing scoring categories by zscores}
table_zscores <- data.frame(
  
  Zscore = c("z ≤ -1.645", 
             "-1.645 < z ≤ -0.84",
             "-0.84 ≤ z ≤ 0.84",
             "0.84 ≤ z ≤ 1.645",
             "z ≥ 1.645"),
  Category = c("Bottom 5% - Lowest",
               "5-20% - Below average",
               "20-80% - Average",
               "80-95% - Above average",
               "Top 5% - Highest") 
)


# Print the table
table_zscores %>%
  kbl() %>%
  kable_minimal(full_width = F)

```


### Assessment Scoring Reliability 

A brief quality control assessment will included for each of the tasks below. This will indicate if the participant had missing data for an item, which can bias their results. In these cases the interpretation of their score will have to be done so with caution. 


### Mental Health (PSC and ASEBA)

Mental health was measured by the **Pediatric Symptom Checklist (PSC)** and the age specific **ASEBA Child Behavior Checklist (CBC)**. The **PSC** is a test that measures a child's emotional, behavioral and psychosocial functioning. It does this by asking questions that require "never", "sometimes", and "always" as responses. These responses then get added to produce a score, in which the higher the score, the higher the emotional or behavioral difficulties are present at a stronger severity. 

```{r reporting PSC }
# Select PSC
PSC <- data %>%
  select(PSC_Performance_z) %>%
  stack()

# Make it prettier
PSC2 <- PSC %>%
  select(Measure = ind, Zscores = values) %>%
  mutate(Zscores = round(Zscores,2))

# Include interpretation
PSC2$Interpretation <- sapply(PSC2$Zscores, zscore_fun)

# Print out the results
PSC2 %>%
  kbl(caption = "") %>%
  kable_styling(full_width = F)

```

```{r PSC violation}
# The violation variable makes no sense for this task- only the NA variable does

if(data$PSC_NA_num == 0) {
  
  paste0("There are no items missing for this score, thus this score is accurate")
  
} else {
  
  paste0("There are items missing for this score, thus this score may be inaccurate")
  
}
```

There are two surveys for the **CBC** that are administered depending on the age of the child. There is one for younger children CBC_3_6 and another for older children CBC_6_18. Similar to the PSC, these items also measure emotional, behavioral and social problems and have "not true", "somewhat or sometimes true", and "very true or often true" responses. Additionally, the higher the score in these measures, the more difficulty the child has with these emotional and behavioral domains. 

Here are some definitions of what the different subscales mean:

- Emotionally reactive: Easily upset, sensitive, mood swings.
- Anxiety/Depression: Excessive worry, sadness, self-consciousness.
- Withdrawn/Depressed: Avoids social interaction, lack of motivation.
- Somatic Complaints: Unexplained physical issues (e.g., headaches, nausea).
- Social problems: Trouble making friends, feeling lonely, being teased.
- Thought problems: Strange ideas, hallucinations, obsessions, odd behaviors.
- Attention problems: Easily distracted, trouble focusing, impulsivity (linked to ADHD).
- Aggressive Behavior: Fighting, arguing, anger issues.
- Rule breaking behavior: Lying, stealing, risky behavior.
- **Internalizing problems (general)**: Issues directed inward  (ex: anxiety/depression/somatic)
- **Externalizing problems (general)**: Issues of outward directed behaviors (ex: aggressive/rule-breaking)
- **Total problems**: The sum of all issues, representing overall difficulties.


```{r reporting ASEBA}
# An index for NAs in CBC3_6 data; A high NA value means the child did the CBC6_18
CBC_Criterium <- rowSums(is.na(select(data,CBC3_6_emotionally_reactive_z:CBC3_6_opositional_defiant_problems_z)))

if (CBC_Criterium <= 4) {
  mental <- data %>%
    select(CBC3_6_emotionally_reactive_z:CBC3_6_opositional_defiant_problems_z)
  
  } else {
    mental <- data %>%
    select(CBC6_18_anxious_depressed_z:CBC6_18_Total_Prob_z)
  
  }

# Convert the data into long format
mental_stacked <- stack(mental)

# Let's do some data cleaning
mental_stacked2 <- mental_stacked %>%
  select(Measure = ind, Zscores = values) %>%
  mutate(Zscores = round(Zscores,2))

# Include interpretation
mental_stacked2$Interpretation <- sapply(mental_stacked2$Zscores, zscore_fun)

# Print out the results
mental_stacked2 %>%
  kbl(caption = "") %>%
  kable_styling(full_width = F)
```


```{r CBC violation checklist}
# Same criterium as above!
if (CBC_Criterium <= 4) {
  CBC3_6_missing <- data$CBC3_6_NA_Num
  
  if (CBC3_6_missing == 0) {
    paste0("CBC is reliable since there is no missing data for any item")
  } else {
    paste0("CBC may be unreliable since there is missing data for ", CBC3_6_missing, " items")
  }
  
  
} else {
  
  CBC6_18_missing <- data$CBC6_18_NA_Num
  
  if (CBC6_18_missing == 0) {
    paste0("CBC is reliable since there is no missing data for any item")
  } else {
    paste0("CBC may be unreliable since there is missing data for ", CBC6_18_missing, " items")
  }
  
}

```
### Receptive Vocabulary

**Receptive vocabulary** refers to a person’s ability to understand words they hear or read. It reflects how well someone can comprehend spoken or written language, even if they don’t actively use those words in their own speech or writing. High scores indicate the child is able to understand spoken words well while lower scores indicate difficulty with this. Lower scores could indicate struggling with following directions, reading comprehension, or general communication. Low scores could also indicate **language delay**, **learning disability**, or **hearing issues**.

```{r reporting on receptive vocabulary}
# Select receptive vocabulary
RecVoc <- data %>%
  select(RV_Performance_z) %>%
  stack()

# Make it prettier
RecVoc2 <- RecVoc %>%
  select(Measure = ind, Zscores = values) %>%
  mutate(Zscores = round(Zscores,2))

# Include interpretation
RecVoc2$Interpretation <- sapply(RecVoc2$Zscores, zscore_fun)

# Print out the results
RecVoc2 %>%
  kbl(caption = "") %>%
  kable_styling(full_width = F)
```

```{r recep vocab violation}
# SR Violation does not apply to this dataset
if(data$RV_NA_num == 0) {
  
  paste0("There are no items missing for this score, thus this score is accurate")
  
} else {
  
  paste0("There are items missing for this score, thus this score may be inaccurate")
  
}
```

### General Cognitive Ability (Reasoning and Memory)

Three assessments were used to capture cognitive ability in the child. These measures are **Pattern Reasoning (PR)**, **Triangles (TR)**, and **Letter & Digit Span (LetDig)**. Scores for each of these measures indicate how well they performed- with higher scores representing better performance. 

- **Pattern Reasoning**: It is an index of fluid intelligence- requires the child to recognize patterns, solve problems, and think logically without requiring experiences to guide them.

- **Triangles**: It is an index of visual spatial processing. It Assesses the ability to mentally manipulate shapes and recognize geometric patterns. 

- **Letter & Digit Span**; A common measure of working memory. It measures the ability to hold and manipulate information in short-term memory. It is essentially for learning and problem solving. 


```{r scoring IQ measures}
# Select the three IQ measures
IQ_measures <- data %>%
  select(PR_Performance_z:LetDig_Performance_z) %>%
  stack()

# Make it prettier
IQ_measures2 <- IQ_measures %>%
  select(Measure = ind, Zscores = values) %>%
  mutate(Zscores = round(Zscores,2))

# Include interpretation
IQ_measures2$Interpretation <- sapply(IQ_measures2$Zscores, zscore_fun)

# Print out the results
IQ_measures2 %>%
  kbl(caption = "") %>%
  kable_styling(full_width = F)
```


```{r IQ measures violation}
full_data$TR_SR_Viola
full_data$PR_SR_Viola
full_data$LetDig_NA.Num
```