---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, echo = F}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)
knitr::opts_chunk$set(warning = FALSE)

```

```{r loading in the packages, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(readxl)
library(kableExtra)
```

```{r pick the ID number}
ID_num = 62 #41 #62 

```


```{r loading in all required data}
# Set working directory to where your GitHub repository is
DataLocation = "~/Documents/Github/KBB/"

# Set working directory to where the final data is
setwd('C:/Users/lledesma.TIMES/Documents/KBB/Data/FINAL_DS/Behavioral')

# Load in the full data
full_data <- read_excel("full_data.xlsx")

# drop any NAs in the dataset for Child_ID
full_data <- drop_na(full_data, Child_ID)
```

```{r load in dd status data}
# Set the location for where the ID tracker is saved and the other where the CFM data is saved
DataLocation2 = "C:/Users/lledesma.TIMES/Documents/KBB/Data/FINAL_DS/Screener/Matched_Siblings/"
DataLocation3 = "C:/Users/lledesma.TIMES/Documents/KBB/Data/FINAL_DS/Screener/"

# Load tracker ID infor
setwd(DataLocation2)
trackerID <- read_excel("Final_ID_Tracker.xlsx")

# Load the CFM data
setwd(DataLocation3)
CFM <- read_excel("All Children.xlsx")

# Some minor data cleaning
trackerID2 <- select(trackerID, full_child_ID = Child_ID, Child_ID = ID)
CFM2 <- select(CFM, full_child_ID = Child_ID, KBB_CFM_DD_type)

# Merge the datasets by child_ID
CFM_ID <- trackerID2 %>%
  full_join(CFM2, by = "full_child_ID") %>%
  select(-full_child_ID)

# Introdude this info into full data
full_data <- full_data %>%
  full_join(CFM_ID, "Child_ID")
  
# drop NA's in Child_ID
full_data <- drop_na(full_data, Child_ID)
```

```{r create groups by age}
# Arbitrary groups
# 2-4;
# 5-8;
# 9-12; 
# 13-15;
# 16-19 

full_data <- full_data %>%
  mutate(age_group = case_when(
     Age >= 2 & Age < 5 ~ "g3_4",
     Age >= 5 & Age < 9 ~ "g5_8",
     Age >= 9 & Age < 13 ~ "g9_12",
     Age >= 13 & Age < 16 ~ "g13_15",
     Age >= 16 & Age < 19 ~ "g16_19",
     TRUE ~ "ERROR"
  ))

# table(full_data$age_group)

# Drop any ERROR (kids have not have data collected yet)
full_data <- filter(full_data, age_group != "ERROR")
```


```{r obtain zscores for each task for each group}
full_data_z <- full_data %>%
  group_by(age_group) %>%
  mutate(KBB_CFM_DD_type,
         Epilepsy,
         PSC_Performance_z = c(scale(PSC_Performance)),
         CBC3_6_emotionally_reactive_z = c(scale(CBC3_6_emotionally_reactive)),
         CBC3_6_anxious_depressed_z = c(scale(CBC3_6_anxious_depressed)),
         CBC3_6_somatic_complaints_z = c(scale(CBC3_6_somatic_complaints)),
         CBC3_6_withdrawn_z = c(scale(CBC3_6_withdrawn)),
         CBC3_6_sleep_problems_z = c(scale(CBC3_6_sleep_problems)),
         CBC3_6_attention_problems_z = c(scale(CBC3_6_attention_problems)),
         CBC3_6_aggressive_behaviors_z = c(scale(CBC3_6_aggressive_behaviors)),
         CBC3_6_other_problems_z = c(scale(CBC3_6_other_problems)),
         CBC3_6_Internalizing_z = c(scale(CBC3_6_Internalizing)),
         CBC3_6_Externalizing_z = c(scale(CBC3_6_Externalizing)),
         CBC3_6_Total_Prob_z = c(scale(CBC3_6_Total_Prob)),
         CBC3_6_depressive_problems_z = c(scale(CBC3_6_depressive_problems)),
         CBC3_6_anxiety_problems_z = c(scale(CBC3_6_anxiety_problems)),
         CBC3_6_autism_spectrum_problems_z = c(scale(CBC3_6_autism_spectrum_problems)),
         CBC3_6_attention_deficit_hyperactivity_problems_z = c(scale(CBC3_6_attention_deficit_hyperactivity_problems)),
         CBC3_6_opositional_defiant_problems_z = c(scale(CBC3_6_opositional_defiant_problems)),
         CBC6_18_anxious_depressed_z = c(scale(CBC6_18_anxious_depressed)),
         CBC6_18_withdrawn_depressed_z = c(scale(CBC6_18_withdrawn_depressed)),
         CBC6_18_somatic_complaints_z = c(scale(CBC6_18_somatic_complaints)),
         CBC6_18_social_problems_z = c(scale(CBC6_18_social_problems)),
         CBC6_18_thought_problems_z = c(scale(CBC6_18_thought_problems)),
         CBC6_18_attention_problems_z = c(scale(CBC6_18_attention_problems)),
         CBC6_18_ruleBreaking_behavior_z = c(scale(CBC6_18_ruleBreaking_behavior)),
         CBC6_18_aggressive_behavior_z = c(scale(CBC6_18_aggressive_behavior)),
         CBC6_18_other_problems_z = c(scale(CBC6_18_other_problems)),
         CBC6_18_Internalizing_z = c(scale(CBC6_18_Internalizing)),
         CBC6_18_Externalizing_z = c(scale(CBC6_18_Externalizing)),
         CBC6_18_Total_Prob_z = c(scale(CBC6_18_Total_Prob)),
         RV_Performance_z = c(scale(RV_Performance)),
         PR_Performance_z = c(scale(PR_Performance)),
         TR_Performance_z = c(scale(TR_Performance)),
         LetDig_Performance_z = c(scale(LetDig_Performance))) %>%
  ungroup()

```

### The KBB Project

```{r, demographic information as Mei suggested,  results='asis', echo = FALSE}
data <- full_data_z %>%
  filter(Child_ID == ID_num)


cat('\n - Individual report for ID:', ID_num, '\n')

cat('\n - Age:',data$Age,'\n')
cat('\n - Sex:',data$Sex,'\n')

```



```{r create a function to report zscore results}
# We decided on the following categories
# Bottom 5% → "Lowest individuals."
# 5% to 20% → "Below average individuals."
# 20% to 80% → "Average individuals."
# 80% to 95% → "Above average individuals."
# Top 5% → "Highest individuals."

zscore_fun <- function(x) {
  if (is.na(x)) {
    return(NA_character_)  # Ensures NA is returned as a character
  }
  
  if (x >= 1.645) {
    return("Top 5% - Highest")
  } else if (x >= 0.84) {
    return("80-95% - Above average")
  } else if (x >= -0.84) {
    return("20-80% - Average")
  } else if (x >= -1.645) {
    return("5-20% - Below average")
  } else {
    return("Bottom 5% - Lowest")
  }
}


```

**Possible Areas of Developmental Disability According to the Screener**

```{r reporting potential difficulties in pretty font,  results='asis', echo = FALSE}
# Report the CFM domains that at least some difficulty was reported
Num_Diff <- data.frame(Domains = do.call(c, str_split(data$KBB_CFM_DD_type, "; ")))

if(sum(is.na(Num_Diff)) == 1) {
  
  cat('\n - No difficulties were reported for the CFM \n')
  
} else {
  
  Num_Diff %>%
    mutate(`Reported Difficulties` = "Yes") %>%
    kbl() %>%
    kable_minimal(full_width = F)
  
}

if(data$Epilepsy == "Yes") {
  cat('\n - A potential risk for having epilepsy \n')
} else {
  cat('\n - No potential risk for having epilepsy \n')
  
}



```




#### 1. Eyesight

The score for seeing ability Is expressed as a fraction. The top number is the distance at which your child can see a letter or symbol. The lower number is the distance at which the average child can see a letter or symbol. Larger values of this lower number indicate the child’s poorer vision.

Summary Examples:

- 20/16 = better than normal vision
- 20/20 = normal vision
- 20/25 = slightly worse than normal, but still functional
- 20/40 = acceptable for daily activities, but not ideal
- 20/100 = significant vision impairment
- 20/200 and worse = legally blind or very poor vision

```{r physical data eyes}
Eyes <- data %>%
  select(Left.eye:Both.Eyes)

# Print the tables
Eyes %>%
  kbl() %>%
  kable_styling(full_width = F)

```

#### 2. Hearing

The score for hearing ability is expressed as decibels (dB; “loudness”) of a sound needed for the child to hear the sound; larger values indicate worse hearing.

Summary Examples:

- -10 to -1 - better than normal hearing:
- 0 to 20 dB - normal hearing:
- 21 to 40 dB - mild hearing loss: may have trouble hearing soft sounds like whispers
- 41 to 70 dB - moderate hearing loss: may have difficulty hearing normal conversations
- 71 to 90 dB - severe hearing loss: may only hear loud speech or sound
- 91 dB+ - profound hearing loss: may not be able to hear everyday sounds without hearing aids
  

```{r physical data hearing}
Ears_L <- data %>%
  select(`1000Hz_dB` = `1000Hz_Left_dB`,
         `2000Hz_dB` = `2000Hz_Left_dB`,
         `4000Hz_dB` = `4000Hz_Left_dB`,) %>%
  mutate(Ear = "Left")

Ears_R <- data %>%
  select(`1000Hz_dB` = `1000Hz_Right_dB`,
         `2000Hz_dB` = `2000Hz_Right_dB`,
         `4000Hz_dB` = `4000Hz_Right_dB`) %>%
  mutate(Ear = "Right")

Ears <- rbind(Ears_L, Ears_R) %>% select(Ear, everything())


Ears %>%
  kbl(caption = "") %>%
  kable_styling(full_width = F)
```

### 3. Assessment

Scores for each child on these tests are calculated to compare the child’s performance with the performance of other children in the study who are in the same age group. Below are the numbers of children in each age group used to calculate your child’s score.


```{r, how many children in the group of comparison,  results='asis', echo = FALSE}
# Extract all children within the current age group of the child
current_age_group_all_data <- full_data %>%
  filter(age_group == data$age_group) 

# Keeps vars of interest that are not filled with NA's
current_age_group_all_data2 <- current_age_group_all_data %>%
  select(RV_StopRule_Num:LetDig_Performance)

cat('\n - There are at most', nrow(current_age_group_all_data2), 'children in this age group for comparison \n')

cat('\n - The score (called a z-score) indicates how far your child’s score is from their age group’s average score.\n' )
```



#### a. Mental Health - emotional, behavioral, and social problems

The PSC measures a child’s general emotional, behavioral and psychosocial functioning. A higher score indicates a higher level of emotional or behavioral difficulties.    

```{r reporting PSC }
# Select PSC
PSC <- data %>%
  select(PSC_Performance_z) %>%
  stack()

# Make it prettier
PSC2 <- PSC %>%
  select(Measure = ind, Zscores = values) %>%
  mutate(Zscores = round(Zscores,2))

# Include interpretation
PSC2$Interpretation <- sapply(PSC2$Zscores, zscore_fun)

# Print out the results
PSC2 %>%
  kbl(caption = "") %>%
  kable_styling(full_width = F)

```

```{r PSC violation}
# The violation variable makes no sense for this task- only the NA variable does

if(data$PSC_NA_num == 0) {
  
  paste0("There are no items missing for this score, thus this score is accurate")
  
} else {
  
  paste0("There are items missing for this score, thus this score may be inaccurate")
  
}
```

There are two surveys for the **CBC** that are administered depending on the age of the child. There is one for younger children CBC_3_6 and another for older children CBC_6_18. Similar to the PSC, these items also measure emotional, behavioral and social problems and have "not true", "somewhat or sometimes true", and "very true or often true" responses. Additionally, the higher the score in these measures, the more difficulty the child has with these emotional and behavioral domains. 

Here are some definitions of what the different subscales mean:

- Emotionally reactive: Easily upset, sensitive, mood swings.
- Anxiety/Depression: Excessive worry, sadness, self-consciousness.
- Withdrawn/Depressed: Avoids social interaction, lack of motivation.
- Somatic Complaints: Unexplained physical issues (e.g., headaches, nausea).
- Social problems: Trouble making friends, feeling lonely, being teased.
- Thought problems: Strange ideas, hallucinations, obsessions, odd behaviors.
- Attention problems: Easily distracted, trouble focusing, impulsivity (linked to ADHD).
- Aggressive Behavior: Fighting, arguing, anger issues.
- Rule breaking behavior: Lying, stealing, risky behavior.
- **Internalizing problems (general)**: Issues directed inward  (ex: anxiety/depression/somatic)
- **Externalizing problems (general)**: Issues of outward directed behaviors (ex: aggressive/rule-breaking)
- **Total problems**: The sum of all issues, representing overall difficulties.


```{r reporting ASEBA}
# An index for NAs in CBC3_6 data; A high NA value means the child did the CBC6_18
CBC_Criterium <- rowSums(is.na(select(data,CBC3_6_emotionally_reactive_z:CBC3_6_opositional_defiant_problems_z)))

if (CBC_Criterium <= 4) {
  mental <- data %>%
    select(CBC3_6_emotionally_reactive_z:CBC3_6_opositional_defiant_problems_z)
  
  } else {
    mental <- data %>%
    select(CBC6_18_anxious_depressed_z:CBC6_18_Total_Prob_z)
  
  }

# Convert the data into long format
mental_stacked <- stack(mental)

# Let's do some data cleaning
mental_stacked2 <- mental_stacked %>%
  select(Measure = ind, Zscores = values) %>%
  mutate(Zscores = round(Zscores,2))

# Include interpretation
mental_stacked2$Interpretation <- sapply(mental_stacked2$Zscores, zscore_fun)

# Print out the results
mental_stacked2 %>%
  kbl(caption = "") %>%
  kable_styling(full_width = F)
```


```{r CBC violation checklist}
# Same criterium as above!
if (CBC_Criterium <= 4) {
  CBC3_6_missing <- data$CBC3_6_NA_Num
  
  if (CBC3_6_missing == 0) {
    paste0("CBC is reliable since there is no missing data for any item")
  } else {
    paste0("CBC may be unreliable since there is missing data for ", CBC3_6_missing, " items")
  }
  
  
} else {
  
  CBC6_18_missing <- data$CBC6_18_NA_Num
  
  if (CBC6_18_missing == 0) {
    paste0("CBC is reliable since there is no missing data for any item")
  } else {
    paste0("CBC may be unreliable since there is missing data for ", CBC6_18_missing, " items")
  }
  
}

```
### Receptive Vocabulary

**Receptive vocabulary** refers to a person’s ability to understand words they hear or read. It reflects how well someone can comprehend spoken or written language, even if they don’t actively use those words in their own speech or writing. High scores indicate the child is able to understand spoken words well while lower scores indicate difficulty with this. Lower scores could indicate struggling with following directions, reading comprehension, or general communication. Low scores could also indicate **language delay**, **learning disability**, or **hearing issues**.

```{r reporting on receptive vocabulary}
# Select receptive vocabulary
RecVoc <- data %>%
  select(RV_Performance_z) %>%
  stack()

# Make it prettier
RecVoc2 <- RecVoc %>%
  select(Measure = ind, Zscores = values) %>%
  mutate(Zscores = round(Zscores,2))

# Include interpretation
RecVoc2$Interpretation <- sapply(RecVoc2$Zscores, zscore_fun)

# Print out the results
RecVoc2 %>%
  kbl(caption = "") %>%
  kable_styling(full_width = F)
```

```{r recep vocab violation}
# SR Violation does not apply to this dataset
if(data$RV_NA_num == 0) {
  
  paste0("There are no items missing for this score, thus this score is accurate")
  
} else {
  
  paste0("There are items missing for this score, thus this score may be inaccurate")
  
}
```

### General Cognitive Ability (Reasoning and Memory)

Three assessments were used to capture cognitive ability in the child. These measures are **Pattern Reasoning (PR)**, **Triangles (TR)**, and **Letter & Digit Span (LetDig)**. Scores for each of these measures indicate how well they performed- with higher scores representing better performance. 

- **Pattern Reasoning**: It is an index of fluid intelligence- requires the child to recognize patterns, solve problems, and think logically without requiring experiences to guide them.

- **Triangles**: It is an index of visual spatial processing. It Assesses the ability to mentally manipulate shapes and recognize geometric patterns. 

- **Letter & Digit Span**; A common measure of working memory. It measures the ability to hold and manipulate information in short-term memory. It is essentially for learning and problem solving. 


```{r scoring IQ measures}
# Select the three IQ measures
IQ_measures <- data %>%
  select(PR_Performance_z:LetDig_Performance_z) %>%
  stack()

# Make it prettier
IQ_measures2 <- IQ_measures %>%
  select(Measure = ind, Zscores = values) %>%
  mutate(Zscores = round(Zscores,2))

# Include interpretation
IQ_measures2$Interpretation <- sapply(IQ_measures2$Zscores, zscore_fun)

# Print out the results
IQ_measures2 %>%
  kbl(caption = "") %>%
  kable_styling(full_width = F)
```


```{r IQ measures violation}
#full_data$PR_SR_Viola
#full_data$TR_SR_Viola
#full_data$LetDig_NA.Num

```