---
title: "KBB Biweekly Screener Report"
output: html_document
---

```{r setup, echo = F}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)

```


```{r loading in the packages, warning = FALSE, echo = F}
library(tidyverse)
library(ggplot2)
library(readxl)
library(kableExtra)
library(openxlsx) # To save excel files with multiple tabs
```

```{r load in the data, echo = F, warning = FALSE}
# Set the working directory
setwd("C:/Users/lledesma.TIMES/Documents/KBB/Data/FINAL_DS/Screener")

# Load the ID data
Comprehensive.Screener.data <- read_excel("Comprehensive Screener Scoring.xlsx")


# Load in the binded data
Binded.data <- read_excel("All Children.xlsx")

# Keep variables of interest 
Binded.data2 <- Binded.data %>%
  select(Child_ID, KBB_CFM_DD_type, Screener.Type) %>%
  mutate(KBB_CFM_DD_status = ifelse(is.na(KBB_CFM_DD_type),"No","Yes"))

# Merge the datasets together
Comprehensive.Screener.data <- Comprehensive.Screener.data %>%
  left_join(Binded.data2, by = "Child_ID")
         
```

### Overview

```{r asdsad, results='asis',echo = FALSE}

cat("This report was created on ",paste(Sys.Date()),".\n\n", sep ="")


cat("The most recent CFM screener was collected on ",last(sort(Comprehensive.Screener.data$Date_of_Evaluation)),".\n\n",sep="") 

cat(nrow(Comprehensive.Screener.data)," CFM and epilepsy screeners have been collected since ",first(sort(Comprehensive.Screener.data$Date_of_Evaluation))," from ",length(unique(Comprehensive.Screener.data$HOH_ID))," households across ",length(unique(Comprehensive.Screener.data$Name_of_the_Village))," villages.", sep = "")


```


```{r Nesting and mapping by overall summary, echo = FALSE, warning = F}
# Nest the data by Overall.Summary
Comprehensive.Screener.data.nested <- Comprehensive.Screener.data %>%
  group_by(Overall.Summary) %>%
  nest()

# Obtain the following information for each group
Comprehensive.Screener.data.nested.summary <- Comprehensive.Screener.data.nested %>%
  mutate(n = map(.x = data, .f = nrow),
         noDD = map(.x = data, .f = ~sum(.x$KBB_DD_status == "No")),
         DD = map(.x = data, .f = ~sum(.x$KBB_DD_status == "Yes")),
         `DD(%)` = round(unlist(DD)/(unlist(DD) + unlist(noDD))*100,1),
         Epilepsy = map(.x = data, .f = ~sum(.x$Epilepsy == "Yes")),
         No.Epilepsy = map(.x = data, .f = ~sum(.x$Epilepsy == "No")),
         `Epilepsy(%)` = round(unlist(Epilepsy)/(unlist(Epilepsy) + unlist(No.Epilepsy))*100,1),
         Male =  map(.x = data, .f = ~sum(.x$Child_Gender == "male")),
         Female = map(.x = data, .f = ~sum(.x$Child_Gender == "female")),
         `Male(%)` = round(unlist(Male)/(unlist(Female) + unlist(Male))*100,1))

Comprehensive.Screener.unnested <- Comprehensive.Screener.data.nested.summary %>%
  select(-data) %>%
  unnest() 

```


```{r creating errors and non errors dataframes, echo = F}
# Save errors separately
errors.vector <- c("Incorrect Screener",
                   "Recruited With Incorrect Screener",
                   "Control for Incorrect Screener",
                   "Should have been Excluded",
                   "Control for Excluded Child",
                   "Manual: Matched Half-Siblings (Old)") 

errors.df <- Comprehensive.Screener.unnested %>%
  filter(Overall.Summary %in% errors.vector) %>%
  select(-c(No.Epilepsy,Female))

errors.df <- errors.df[match(errors.vector,errors.df$Overall.Summary),]

errors.df <- errors.df %>%
  mutate(Overall.Summary = paste(Overall.Summary," (n=",n,")",sep="")) %>%
  select(-n)


# Save the non errors portion
report.vector <- c("Excluded Children",
                   "No Matches Within HOH",
                   "Manual: Unable to Match",
                   "Manual: Matched Half-Siblings",
                   "Manual: Matched Cousins",
                   "Manual: Matched Siblings")

report.df <- Comprehensive.Screener.unnested %>%
  filter(Overall.Summary %in% report.vector) %>%
  select(-c(No.Epilepsy,Female))

report.df <- report.df[match(report.vector,report.df$Overall.Summary),]

report.df <- report.df %>%
  mutate(Overall.Summary = paste(Overall.Summary," (n=",n,")",sep="")) %>%
  select(-n)
```

### Errors Table

The table below describes the number of errors in data collection so far. These children are **not** present in the report table. 

- **Incorrect Screener**: The wrong CFM screener was used on a child. Either a child much younger than 5 years was screened with the CFM5-17 or a children much older than 4 years old was screened with the CFM2-4.
- **Recruited With Incorrect Screener**: Number of children that were recruited and given ID with an incorrect screener. Their data will not be included in behavioral/neuroimaging analyses.
- **Controls for Incorrect Screener**: If a proband is given an ID then so is their sibling. These are the siblings that correspond with the probands receiving the incorrect screener.
- **Should have been Excluded**: These are proband that should have been excluded because they are either blind, deaf, or immobile. However, they were recruited and given an ID. Their data will not be included in behavioral/neuroimaging analyses.
- **Control for Excluded Children**: If a proband is given an ID then so is their sibling. These are the siblings that correspond with the probands that should have been excluded.
- **Manual: Matched Half-Siblings (Old):** These are half-sibling matches that were given an ID too late- they reside in an area where the team will no longer be collecting data from- thus they will a part of the error category. Note: Their originally given ID's have been removed and given to other children. 

```{r print out the error table, echo = FALSE}
errors.df %>% 
  kbl() %>%
  kable_paper(full_width = F)

```

These errors (except for the manual half-silbing (old)) occurred before code was created to identify the children and prevent them from being recruited. This code is now active thus only the 'Incorrect Screener' number should change from now on. 

### Report Table

The numbers reported in this table and those above are **mutually exclusive**. No child in one group will be present in another, this is both true across these two datasets and within them (This applies for the groups expressed in 'Overall.Summary'. 

- **Excluded Chidren**: A proband who has "Cannot at all" reported for seeing, hearing, or walking. A non-DD sibling with at least "some difficulty" in any of these domains. 
- **No Matches Within HOH**: All the children that were in a household with all DD or all noDD, thus there was no way to match between siblings.
- **Manual: Unable to Match**: Children within a household that potentially was a DD and nonDD pair, however, they were not recruited. Either due to not having another pair available in the household (Ex: 1 DD and 3 noDD siblings) or because two non-related children are living in the same HOH.
- **Manual: Matched Half-Siblings**: Matched half-sibling pairs (Given IDs; Began March 21, 2025)
- **Manual: Matched Cousins**: Matched cousin pairs (Given IDs; Began May 8th, 2025)
- **Manual: Matched Siblings**: Matched sibling pairs that have been recruited (Given IDs)


```{r print our the main table, echo = FALSE}
report.df %>% 
  kbl() %>%
  kable_paper(full_width = F)

```

### Correct Screener Compiled

- These are all the numbers so far not including:
- a) 'Incorrect Screener'
- b) 'Recruited With Incorrect Screener'

```{r create a table for a correct compiled screener, echo = FALSE, warning = FALSE}
incorrect.vector <- c("Incorrect Screener",
                      "Recruited With Incorrect Screener")

Correct.Comprehensive.Screener.data <- Comprehensive.Screener.data %>%
  filter(!(Overall.Summary %in% incorrect.vector))

Correct.Comprehensive.Screener.data.nested <- nest(Correct.Comprehensive.Screener.data)

Correct.Comprehensive.Screener.data.nested.summary <- Correct.Comprehensive.Screener.data.nested %>%
  mutate(n = map(.x = data, .f = nrow),
         noDD = map(.x = data, .f = ~sum(.x$KBB_DD_status == "No")),
         DD = map(.x = data, .f = ~sum(.x$KBB_DD_status == "Yes")),
         `DD(%)` = round(unlist(DD)/(unlist(DD) + unlist(noDD))*100,1),
         Epilepsy = map(.x = data, .f = ~sum(.x$Epilepsy == "Yes")),
         No.Epilepsy = map(.x = data, .f = ~sum(.x$Epilepsy == "No")),
         `Epilepsy(%)` = round(unlist(Epilepsy)/(unlist(Epilepsy) + unlist(No.Epilepsy))*100,1),
         Male =  map(.x = data, .f = ~sum(.x$Child_Gender == "male")),
         Female = map(.x = data, .f = ~sum(.x$Child_Gender == "female")),
         `Male(%)` = round(unlist(Male)/(unlist(Female) + unlist(Male))*100,1))

Correct.Comprehensive.Screener.unnested <- Correct.Comprehensive.Screener.data.nested.summary %>%
  select(-data) %>%
  unnest()

Comprehensive.Screener.df <- Correct.Comprehensive.Screener.unnested %>%
  select(-c(No.Epilepsy, Female))

# Print the Table
Comprehensive.Screener.df %>% 
  kbl() %>%
  kable_paper(full_width = F)

```

**Disclaimer**: Please note that from all of these screeners we have this number of children with epilepsy. However, there is technically another child with epilepsy not present here and that is because they were screened with the wrong CFM type. 


### CFM DD and Epilepsy Breakdown by Sex

```{r create a DD sample breakdown, echo = FALSE}
Comprehensive.Screener.data <- Comprehensive.Screener.data %>%
  mutate(KBB_DD_status_Broken = case_when(
        KBB_CFM_DD_status == "Yes" & Epilepsy == "Yes" ~ "CFM and Epilepsy",
        KBB_CFM_DD_status == "Yes" ~ "CFM Only",
        Epilepsy == "Yes" ~ "Epilepsy Only",
        TRUE ~ "NoDD"
    
  ) 
  
)

# Remove incorrect screeners
Comprehensive.Screener.data <- filter(Comprehensive.Screener.data, !(Overall.Summary %in% c("Incorrect Screener", "Recruited With Incorrect Screener")))

# Create the table of interest
DD.Epilepsy.Sex.df <- data.frame(cbind(table(Comprehensive.Screener.data$KBB_DD_status_Broken, Comprehensive.Screener.data$Child_Gender))) %>%
  mutate(total = female + male)
  
# Add a row for total
DD.Epilepsy.Sex.df <- rbind(DD.Epilepsy.Sex.df, c(colSums(DD.Epilepsy.Sex.df)))
row.names(DD.Epilepsy.Sex.df)[5] <- "Total"

# Print the table
DD.Epilepsy.Sex.df %>%
  kbl(caption = "Frequency of CFM DD and Epilepsy by Sex") %>%
  kable_paper(full_width = F)
    
```


### Types of CFM DD (Includes CFM and Epilepsy DD)

Below are tables that describe the prevalence and frequency of different types of DD for each sex based on the screener used (CFM2_4 vs CFM5_17). From this information 4 tables are made, one for each sex for each screener type. 

- 'Freq' = The total number of reported difficulties within that group (not mutually exclusive)

- 'N' =  Indicates the total number of children that have DD within that group.

- 'Percentage' = Shows the percentage of the reported frequency within that group

- 'Group' = Used as quality control to show which group these numbers belong to. 

```{r types of CFM DD, echo = F}
# Save all rows with reported difficulties
CFM.DD <- Comprehensive.Screener.data %>%
  filter(KBB_DD_status_Broken %in% c("CFM and Epilepsy", "CFM Only"))


CFM.DD.summarize <- CFM.DD %>%
  group_by(Screener.Type, Child_Gender) %>%
  summarise(Difficulty.num = length(KBB_CFM_DD_type),
            Difficulty.types = list(do.call(c,str_split(KBB_CFM_DD_type, pattern = "; "))),
            Table.difficulty.types = list(data.frame(table(Difficulty.types))))


# Export the table information
Table.Report <- list()

for(ii in 1:nrow(CFM.DD.summarize)) {

  group.name <- paste(CFM.DD.summarize$Screener.Type[ii],": " , CFM.DD.summarize$Child_Gender[ii],sep = "")
  current.freq.table <- CFM.DD.summarize$Table.difficulty.types[ii] %>% data.frame()
  current.freq.table$N <- CFM.DD.summarize$Difficulty.num[ii]
  current.freq.table$Percentage <- paste(round(current.freq.table$Freq/ CFM.DD.summarize$Difficulty.num[ii],3)*100,"%", sep ="")
  current.freq.table$Group <- group.name
  current.freq.table <- arrange(current.freq.table, Freq)
  
  Table.Report[[ii]] <- current.freq.table
}

Table.Report[[1]] %>%
  kbl(caption = "Frequency of DD by type (CFM2_4 Females)") %>%
  kable_paper(full_width = F)

Table.Report[[2]] %>%
  kbl(caption = "Frequency of DD by type (CFM2_4 Males)") %>%
  kable_paper(full_width = F)

Table.Report[[3]] %>%
  kbl(caption = "Frequency of DD by type (CFM5_17 Females)") %>%
  kable_paper(full_width = F)

Table.Report[[4]] %>%
  kbl(caption = "Frequency of DD by type (CFM5_17 Males)") %>%
  kable_paper(full_width = F)
```




