---
title: "KBB NIHRPPAnnualReport"
author: "Leandro Ledesma"
date: "2025-05-22"
output: html_document
---

```{r setup, echo = F}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)
knitr::opts_chunk$set(warning = FALSE)

```

```{r loading in the packages, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(readxl)
library(kableExtra)
library(psych)
```

```{r load in the data and merging}
# Set the working directory
setwd("C:/Users/lledesma.TIMES/Documents/KBB/Data/FINAL_DS/Screener")

# Load in the binded data
Binded.data <- read_excel("All Children.xlsx")

# Set working directory again
setwd("~/KBB/Data/FINAL_DS/Screener/Matched_Siblings")

# Load the recruited children (children given IDs)
dat <- read_excel("Final_ID_Tracker.xlsx")

# Keep variables of interest 
Binded.data2 <- Binded.data %>%
  select(Child_ID, KBB_CFM_DD_type, Screener.Type) %>%
  mutate(KBB_CFM_DD_status = ifelse(is.na(KBB_CFM_DD_type),"No","Yes"))

# Merge the datasets together
dat2 <- dat %>%
  left_join(Binded.data2, by = "Child_ID")
```

Here we will be running some descriptives on our dataset to update the tables for the NIHRPPRAnnualReport. The dataset chosen is from the Final_ID_tracker since that contains information from the recruited children. However, to update the numbers from Figure 1, use the bi-weekly report!!

From the IDs that were given, we need to remove any erroneous ones from the dataset

```{r remove bad IDS, include = T}
dat2 <- dat2 %>%
  filter(!Overall.Summary %in% c("Control for Excluded Child",
                                "Control for Incorrect Screener",
                                "Recruited With Incorrect Screener",
                                "Should have been Excluded"))

```


Qaulity Control Measures
```{r sample breakdown by relationship type}
paste0("There are ", nrow(dat2), " children in this dataset with IDs- aka recruited children")

addmargins(table(dat2$Overall.Summary)) %>%
  kbl(caption = "Recruited children by paired group") %>%
  kable_minimal(full_width = F)


table(dat2$KBB_DD_status) %>%
  kbl(caption = "Total DD vs non-DD") %>%
  kable_minimal(full_width = F)


table(dat2$Overall.Summary, dat2$KBB_DD_status) %>%
  kbl(caption = "Total DD vs non-DD by group pari") %>%
  kable_minimal(full_width = F)

```



```{r producing first part of table 2}
rbind(describe(dat2$Child_age), 
do.call(rbind, describeBy(dat2$Child_age, dat2$Child_Gender))) %>%
  round(1) %>%
  data.frame() %>%
  transmute(n, 
            Age_range = paste0(min,"-",max), 
            Mea_age_sd = paste0(mean,"(",sd,")")) %>%
  select(N = n, Age_range, Mea_age_sd) %>%
  kbl(caption = "All recruited children count + Age") %>%
  kable_minimal(full_width = F)
  
Probands <- filter(dat2, KBB_DD_status == "Yes")

rbind(describe(Probands$Child_age), 
do.call(rbind, describeBy(Probands$Child_age, Probands$Child_Gender))) %>%
  round(1) %>%
  data.frame() %>%
  transmute(n, 
            Age_range = paste0(min,"-",max), 
            Mea_age_sd = paste0(mean,"(",sd,")")) %>%
  select(N = n, Age_range, Mea_age_sd) %>%
  kbl(caption = "All proband children (including epilspey only) count + Age") %>%
  kable_minimal(full_width = F)


```

This next section is for the left side of the second part of table 2 (CFM Proband only)
```{r the age descriptives by screener type}
CFM2_4 <- filter(dat2, Screener.Type == "CFM2_4" & KBB_CFM_DD_status == "Yes")
CFM5_18 <- filter(dat2, Screener.Type == "CFM5_17" & KBB_CFM_DD_status == "Yes")

rbind(describe(CFM2_4$Child_age), 
do.call(rbind, describeBy(CFM2_4$Child_age, CFM2_4$Child_Gender))) %>%
  round(1) %>%
  data.frame() %>%
  transmute(n, 
            Age_range = paste0(min,"-",max), 
            Mea_age_sd = paste0(mean,"(",sd,")")) %>%
  select(N = n, Age_range, Mea_age_sd) %>%
  kbl(caption = "CFM2-4 children count + Age") %>%
  kable_minimal(full_width = F)


rbind(describe(CFM5_18$Child_age), 
do.call(rbind, describeBy(CFM5_18$Child_age, CFM5_18$Child_Gender))) %>%
  round(1) %>%
  data.frame() %>%
  transmute(n, 
            Age_range = paste0(min,"-",max), 
            Mea_age_sd = paste0(mean,"(",sd,")")) %>%
  select(N = n, Age_range, Mea_age_sd) %>%
  kbl(caption = "CFM5-17 children count + Age") %>%
  kable_minimal(full_width = F)


```

This is for the right side of the second half of Table2

```{r frequency of number of difficulties by screener type and sex}
CFM2_4$Diff_count <- sapply(str_split(CFM2_4$KBB_CFM_DD_type, "; "), function(x) length(x))
CFM5_18$Diff_count <- sapply(str_split(CFM5_18$KBB_CFM_DD_type, "; "), function(x) length(x))

rbind(
table(CFM2_4$Diff_count),
t(table(CFM2_4$Diff_count, CFM2_4$Child_Gender))
) %>%
  kbl(caption = "CFM2-4 Number of At least some difficulties") %>%
  kable_minimal(full_width = F)


table(CFM2_4$Child_Gender, CFM2_4$Epilepsy) %>%
  kbl(caption = "CFM2-4 Epilepsy Count") %>%
  kable_minimal(full_width = F)
  

rbind(
table(CFM5_18$Diff_count),
t(table(CFM5_18$Diff_count, CFM5_18$Child_Gender))
) %>%
  kbl(caption = "CFM5_18 Number of At least some difficulties") %>%
  kable_minimal(full_width = F)

```

Epilepsy Count information from recruited children
```{r create tables that inform us of the epilspey counts}
# KBB_CFM_DD_status = means are they DD stricly from CFM screening!
dat2 <- dat2 %>%
  mutate(Epilepsy_CFM = case_when(
    Epilepsy == "Yes" & KBB_CFM_DD_status == "Yes" ~ "CFM and Epilepsy",
    Epilepsy == "No" & KBB_CFM_DD_status == "Yes" ~ "CFM Only",
    Epilepsy == "Yes" & KBB_CFM_DD_status == "No" ~ "Epilepsy Only",
    Epilepsy == "No" & KBB_CFM_DD_status == "No" ~ "NoDD",
    TRUE ~ "Error"
  ))

addmargins(table(dat2$Epilepsy_CFM, dat2$Child_Gender)) %>%
  kbl(caption = "Frequency of CFM DD and Epilepsy by Sex (CFM5-17") %>%
  kable_minimal(full_width = F)

```
