---
title: "Preliminary Power Analyses on Behavioral Data"
author: "Leandro Ledesma"
date: "2025-05-30"
output: html_document
---

```{r setup, echo = F}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)
knitr::opts_chunk$set(warning = FALSE)

```

```{r loading in the packages, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(readxl)
library(psych)
library(lavaan)
library(kableExtra)
library(lme4)
library(lmerTest)
library(effectsize)
library(simr)
```

## Load in the data
- The IDs with DD information
- Demographic information
- The scored ZAT scores
- The scored Vineland-II

```{r load in the data}
# Set working directory
setwd("C:/Users/lledesma.TIMES/Documents/KBB/Data/FINAL_DS")

# Load in the data
IDs <- read_excel("Screener/Matched_Siblings/Final_ID_Tracker.xlsx")
demo <- read_excel("Demographics/Demographics.xlsx")
ZAT <- read_excel("Behavioral/Children/ZAT.xlsx")
VinelandII <- read_excel("Behavioral/Adults/VinelandII.xlsx")
DLS <- read_excel("Behavioral/Children/LettrDig.xlsx")
PR <- read_excel("Behavioral/Children/PatternReas.xlsx")
TR <- read_excel("Behavioral/Children/Triangles.xlsx")
BRIEF_P <-  read_excel("Behavioral/Adults/BRIEF2_Parent.xlsx")

# Drop duplicates from demo if any
demo_dup <- demo$Child_ID[duplicated(demo$Child_ID)]
demo <- filter(demo, !Child_ID %in% demo_dup)

# Drop any NAs from the Child_ID
IDs <- drop_na(IDs, Child_ID)
demo <- drop_na(demo, Child_ID)
ZAT <- drop_na(ZAT, Child_ID)
VinelandII <- drop_na(VinelandII, Child_ID)
DLS <- drop_na(DLS, Child_ID)
PR <- drop_na(PR, Child_ID)
TR <- drop_na(TR, Child_ID)
BRIEF_P <- drop_na(BRIEF_P, Child_ID)

# Some data cleaning
IDs$Child_ID <- as.numeric(IDs$Child_ID)
IDs$HOH_ID_num <- as.numeric(factor(IDs$HOH_ID))
ZAT <- filter(ZAT, !Child_ID %in% c(463106, 6375))
VinelandII <- filter(VinelandII, !Child_ID %in% c(10444))
PR <- filter(PR, !Child_ID %in% c(1472))
BRIEF_P <- filter(BRIEF_P, !Child_ID %in% c(0, 9474))
demo <- mutate(demo, Age_C = c(scale(Age)))


# Select vars of interest for each dataset
IDs <- select(IDs, Child_ID = ID, HOH_ID,HOH_ID_num, KBB_DD_status)
demo <- select(demo, Child_ID, Sex, Age, Age_C)
ZAT <- select(ZAT, Child_ID, scored_RR:scored_P)
VinelandII <- select(VinelandII, Child_ID, scored_CommR:scored_SCS)

# Join the datasets into one for ZAT
full_ZAT <- ZAT %>%
  left_join(demo ,by = "Child_ID") %>%
  left_join(IDs ,by = "Child_ID")

# Join the datasets into one for Vineland-II
full_VinelandII <- VinelandII %>%
  left_join(demo ,by = "Child_ID") %>%
  left_join(IDs ,by = "Child_ID")

# Join the datasets into one for DLS
full_DLS <- DLS %>%
  left_join(demo ,by = "Child_ID") %>%
  left_join(IDs ,by = "Child_ID")

# Join the datasets into one for PR
full_PR <- PR %>%
  left_join(demo ,by = "Child_ID") %>%
  left_join(IDs ,by = "Child_ID")

# Join the datasets into one for TR
full_TR <- TR %>%
  left_join(demo ,by = "Child_ID") %>%
  left_join(IDs ,by = "Child_ID")

# Join the datasets into one for BRIEF_P
full_BRIEF_P <- BRIEF_P %>%
  left_join(demo ,by = "Child_ID") %>%
  left_join(IDs ,by = "Child_ID")


# Generate composite scores
full_ZAT <- mutate(full_ZAT, ZAT_composite = scored_RR + scored_RC + scored_M + scored_P)
full_VinelandII <- mutate(full_VinelandII, VinelandII_composite = scored_CommR + scored_CommE + scored_DLSP + scored_DLSD + scored_DLSC + scored_SI + scored_SPI + scored_SCS)

# Quickly investigate sample size
nrow(full_ZAT)
nrow(full_VinelandII)
nrow(full_DLS)
nrow(full_PR)
nrow(full_TR)
nrow(full_BRIEF_P)

length(unique(full_ZAT$HOH_ID))
length(unique(full_VinelandII$HOH_ID))
length(unique(full_DLS$HOH_ID))
length(unique(full_PR$HOH_ID))
length(unique(full_TR$HOH_ID))
length(unique(full_BRIEF_P$HOH_ID))
```


## Visulize Unadjusted Predictors of the ZAT


```{r visualizing the ZAT Subtest}
# Visualize the effect of DD Status on a Composite score of the ZAT
full_ZAT %>%
  ggplot(aes(x = KBB_DD_status, y = ZAT_composite)) +
  geom_boxplot() +
  geom_jitter(width = .1, size = .75) +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 4, fill = "red") +
  theme_classic() +
  labs(title = "Unadjusted effect of DD Status on ZAT 4 Subtest Composite Score")


# Visualize the effect of DD Status on the 4 ZAT subtests
full_ZAT %>%
  select(Child_ID, KBB_DD_status, scored_RR:scored_P) %>%
  pivot_longer(-c(Child_ID, KBB_DD_status),
               names_to = "Subtests",
               values_to = "Z-Scores") %>%
  mutate(Subtests = gsub("scored_","",Subtests)) %>%
  group_by(Subtests) %>%
  mutate(z_scores = c(scale(`Z-Scores`))) %>%
  ggplot(aes(x = KBB_DD_status, y = z_scores)) +
  facet_grid(~Subtests) +
  geom_violin() +
  geom_jitter(width = .2, size = .5) +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 4, fill = "red") +
  theme_classic() +
  labs(title = "Unadjusted effect of DD Status on each ZAT 4 Subtest Scores")

# Visualize potential effects of household (random effects)
full_ZAT %>%
  group_by(HOH_ID) %>%
  summarize(mean_ZAT_composite = mean(ZAT_composite)) %>%
  filter(!is.na(mean_ZAT_composite)) %>%
  mutate(HOH_ID = forcats::fct_reorder(HOH_ID, mean_ZAT_composite)) %>%
  ggplot(aes(x = mean_ZAT_composite, y = HOH_ID)) +
  geom_point(size = 1) +
  labs(x = "Household Mean Score",
       y = "Household (ordered)",
       title = "Unadjusted Household Means of Child Scores (ZAT)\n(These are 356 HOH)") +
  theme_classic()

# Visualize the Number of Children within each Household
# Step 1: Count children per household by KBB_DD_status
full_ZAT <- full_ZAT %>%
  group_by(HOH_ID) %>%
  mutate(HOH_num_Children = length(Child_ID))

table(full_ZAT$HOH_num_Children, full_ZAT$KBB_DD_status) %>%
  data.frame() %>%
  rename(Number_Children_in_HOH = Var1, KBB_DD_status = Var2, Number_of_Households = Freq) %>%
  ggplot(aes(x = Number_Children_in_HOH, y = Number_of_Households, fill = KBB_DD_status)) +
  stat_summary(fun = "mean", geom = "bar", position = position_dodge2()) +
  theme_classic() +
  labs(title = "Number of Households with Varying Number of Children ZAT data was collected\nfrom (ZAT data is collected from ages 7-18)")

```




## Visulize Unadjusted Predictors of the Vineland-II
```{r visualizing the VinelandII}
full_VinelandII %>%
  ggplot(aes(x = KBB_DD_status, y = VinelandII_composite)) +
  geom_boxplot() +
  geom_jitter(width = .1, size = .75) +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 4, fill = "red") +
  theme_classic() +
  labs(title = "Unadjusted effect of DD Status on the Vineland-II 8 Subtest Composite Score")


full_VinelandII %>%
  select(Child_ID, KBB_DD_status, scored_CommR:scored_SCS) %>%
  pivot_longer(-c(Child_ID, KBB_DD_status),
               names_to = "Subtests",
               values_to = "Sum of Reponses") %>%
  mutate(Subtests = gsub("scored_","",Subtests)) %>%
  ggplot(aes(x = KBB_DD_status, y = `Sum of Reponses`)) +
  facet_grid(~Subtests) +
  geom_boxplot() +
  geom_jitter(width = .2, size = .5) +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 3.5, fill = "red") +
  theme_classic() +
  labs(title = "Unadjusted effect of DD Status on each Vineland-II 8 Subtest Scores")


# Visualize potential effects of household (random effects)
full_VinelandII %>%
  group_by(HOH_ID) %>%
  summarize(mean_VinelandII_composite = mean(VinelandII_composite)) %>%
  filter(!is.na(mean_VinelandII_composite)) %>%
  mutate(HOH_ID = forcats::fct_reorder(HOH_ID, mean_VinelandII_composite)) %>%
  ggplot(aes(x = mean_VinelandII_composite, y = HOH_ID)) +
  geom_point(size = 1) +
  labs(x = "Household Mean Score",
       y = "Household (ordered)",
       title = "Unadjusted Household Means of Child Scores (VinelandII)\n(These are 360 HOH)") +
  theme_classic()

# Visualize the Number of Children within each Household
# Step 1: Count children per household by KBB_DD_status
full_VinelandII <- full_VinelandII %>%
  group_by(HOH_ID) %>%
  mutate(HOH_num_Children = length(Child_ID))

table(full_VinelandII$HOH_num_Children, full_VinelandII$KBB_DD_status) %>%
  data.frame() %>%
  rename(Number_Children_in_HOH = Var1, KBB_DD_status = Var2, Number_of_Households = Freq) %>%
  ggplot(aes(x = Number_Children_in_HOH, y = Number_of_Households, fill = KBB_DD_status)) +
  stat_summary(fun = "mean", geom = "bar", position = position_dodge2()) +
  theme_classic() +
  labs(title = "Number of Households with Varying Number of Children Vineland-II data was\ncollected from (Vineland-II data is collected from 3-18)")
```


## Visulize Unadjusted Predictors of the Digit and Letter Span

```{r visualizing the Digit and Letter Span}
full_DLS %>%
  ggplot(aes(x = KBB_DD_status, y = LetDig_Performance)) +
  geom_boxplot() +
  geom_jitter(width = .1, size = .75) +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 4, fill = "red") +
  theme_classic() +
  labs(title = "Unadjusted effect of DD Status on the Letter & Dig 4 Subtest Composite Score")


full_DLS %>%
  select(Child_ID, KBB_DD_status, NF_Performance, NB_Performance, LF_Performance, LB_Performance) %>%
  pivot_longer(-c(Child_ID, KBB_DD_status),
               names_to = "Subtests",
               values_to = "Sum of Reponses") %>%
  mutate(Subtests = gsub("scored_","",Subtests)) %>%
  ggplot(aes(x = KBB_DD_status, y = `Sum of Reponses`)) +
  facet_grid(~Subtests) +
  geom_boxplot() +
  geom_jitter(width = .2, size = .5) +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 3.5, fill = "red") +
  theme_classic() +
  labs(title = "Unadjusted effect of DD Status on each Letter & Dig 4 Subtest Scores")


# Visualize the Number of Children within each Household
# Step 1: Count children per household by KBB_DD_status
full_DLS <- full_DLS %>%
  group_by(HOH_ID) %>%
  mutate(HOH_num_Children = length(Child_ID))

table(full_DLS$HOH_num_Children, full_DLS$KBB_DD_status) %>%
  data.frame() %>%
  rename(Number_Children_in_HOH = Var1, KBB_DD_status = Var2, Number_of_Households = Freq) %>%
  ggplot(aes(x = Number_Children_in_HOH, y = Number_of_Households, fill = KBB_DD_status)) +
  stat_summary(fun = "mean", geom = "bar", position = position_dodge2()) +
  theme_classic() +
  labs(title = "Number of Households with Varying Number of Children Let & Dig data was\ncollected from (Letter and Dig data is collected from 3-18)")
```

## Visulize Unadjusted Predictors of Pattern Reasoning

```{r visualizing Pattern Reasoning}
full_PR %>%
  ggplot(aes(x = KBB_DD_status, y = PR_Performance)) +
  geom_boxplot() +
  geom_jitter(width = .1, size = .75) +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 4, fill = "red") +
  theme_classic() +
  labs(title = "Unadjusted effect of DD Status on the Pattern Reasoning Performance")


# Visualize the Number of Children within each Household
# Step 1: Count children per household by KBB_DD_status
full_PR <- full_PR %>%
  group_by(HOH_ID) %>%
  mutate(HOH_num_Children = length(Child_ID))

table(full_PR$HOH_num_Children, full_PR$KBB_DD_status) %>%
  data.frame() %>%
  rename(Number_Children_in_HOH = Var1, KBB_DD_status = Var2, Number_of_Households = Freq) %>%
  ggplot(aes(x = Number_Children_in_HOH, y = Number_of_Households, fill = KBB_DD_status)) +
  stat_summary(fun = "mean", geom = "bar", position = position_dodge2()) +
  theme_classic() +
  labs(title = "Number of Households with Varying Number of Children Pattern Reasoning was\ncollected from (Pattern Reasoning data is collected from 5-18)")
```

## Visulize Unadjusted Predictors of Triangles

```{r visualizing Triangles}
full_TR %>%
  ggplot(aes(x = KBB_DD_status, y = TR_Performance)) +
  geom_boxplot() +
  geom_jitter(width = .1, size = .75) +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 4, fill = "red") +
  theme_classic() +
  labs(title = "Unadjusted effect of DD Status on Triangles Performance")


# Visualize the Number of Children within each Household
# Step 1: Count children per household by KBB_DD_status
full_TR <- full_TR %>%
  group_by(HOH_ID) %>%
  mutate(HOH_num_Children = length(Child_ID))

table(full_TR$HOH_num_Children, full_TR$KBB_DD_status) %>%
  data.frame() %>%
  rename(Number_Children_in_HOH = Var1, KBB_DD_status = Var2, Number_of_Households = Freq) %>%
  ggplot(aes(x = Number_Children_in_HOH, y = Number_of_Households, fill = KBB_DD_status)) +
  stat_summary(fun = "mean", geom = "bar", position = position_dodge2()) +
  theme_classic() +
  labs(title = "Number of Households with Varying Number of Children Triangles was\ncollected from (Triangles data is collected from 3-18)")
```

## Visulize Unadjusted Predictors of BRIEF2 Parent Form

```{r visualizing BRIEF2}
full_BRIEF_P %>%
  ggplot(aes(x = KBB_DD_status, y = GEC)) +
  geom_boxplot() +
  geom_jitter(width = .1, size = .75) +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 4, fill = "red") +
  theme_classic() +
  labs(title = "Unadjusted effect of DD Status on BRIEF-2")


full_BRIEF_P %>%
  ungroup() %>%
  select(Child_ID, KBB_DD_status, Emotional_Control:Working_Memory) %>%
  pivot_longer(-c(Child_ID, KBB_DD_status),
               names_to = "Subtests",
               values_to = "Sum of Reponses") %>%
  mutate(Subtests = gsub("scored_","",Subtests)) %>%
  ggplot(aes(x = KBB_DD_status, y = `Sum of Reponses`)) +
  facet_grid(~Subtests) +
  geom_boxplot() +
  geom_jitter(width = .2, size = .5) +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 3.5, fill = "red") +
  theme_classic() +
  labs(title = "Unadjusted effect of DD Status on each BRIEF2 _Parent Scales")


# Visualize the Number of Children within each Household
# Step 1: Count children per household by KBB_DD_status
full_BRIEF_P <- full_BRIEF_P %>%
  group_by(HOH_ID) %>%
  mutate(HOH_num_Children = length(Child_ID))

table(full_BRIEF_P$HOH_num_Children, full_BRIEF_P$KBB_DD_status) %>%
  data.frame() %>%
  rename(Number_Children_in_HOH = Var1, KBB_DD_status = Var2, Number_of_Households = Freq) %>%
  ggplot(aes(x = Number_Children_in_HOH, y = Number_of_Households, fill = KBB_DD_status)) +
  stat_summary(fun = "mean", geom = "bar", position = position_dodge2()) +
  theme_classic() +
  labs(title = "Number of Households with Varying Number of Children BRIEF-2 Parent was\ncollected from (BRIEF-2 Parent data is collected from 5-18)")
```



## Analysis of the ZAT

```{r analyzing the ZAT}
# Run an empty model
m0_zat <- lmer(ZAT_composite ~ (1 | HOH_ID_num), data = full_ZAT)
performance::icc(m0_zat)
summary(m0_zat)

# Run the full model
m1_zat <- lmer(ZAT_composite ~  Sex + Age_C + KBB_DD_status + (1 | HOH_ID_num), data = full_ZAT)
summary(m1_zat)

# Run a model with just KBB_DD_Status
m2_zat <- lmer(ZAT_composite ~  KBB_DD_status + (1 | HOH_ID_num), data = full_ZAT)

# Calculate effect sizes (limitation)
t_to_eta2(-2.57, 404.714) # Weak effect size 

# Run a power analysis
powerSim(m2_zat)

# Extend the model to 540 clusters (adjust as needed)
cluster_n = 540
m3_zat <- extend(m2_zat, along = "HOH_ID_num", n = cluster_n) 

# Run powerCurve with explicit sample sizes using 'breaks'
zat_pc <- powerCurve(m3_zat, along = "HOH_ID_num", breaks = c(360, 380, 400, 420, 440, 460, 480, 500, 520, 540), 
                     nsim = 1000)

# Print the power curve results
print(zat_pc)
plot(zat_pc)
```

## Analysis of the Vineland-II

```{r analyzing the Vineland}
# Run an empty model
m0_vin <- lmer(VinelandII_composite ~ (1 | HOH_ID), data = full_VinelandII)
performance::icc(m0_vin)
summary(m0_vin)

# Run the full model
m1_vin <- lmer(VinelandII_composite ~  Sex + Age_C + KBB_DD_status + (1 | HOH_ID), data = full_VinelandII)
summary(m1_vin)

# Run a model with just KBB_DD_Status
m2_vin <- lmer(VinelandII_composite ~  KBB_DD_status + (1 | HOH_ID), data = full_VinelandII)

# Calculate effect sizes (limitation)
t_to_eta2(1.055, 460.364) # Inaccurate please address 

# Run a power analysis
powerSim(m2_vin)

# Extend the model to 500 clusters (adjust as needed)
cluster_n = 500
m3_vin <- extend(m2_vin, along = "HOH_ID_num", n = cluster_n) 

# Run powerCurve with explicit sample sizes using 'breaks'
vin_pc <- powerCurve(m3_vin, along = "HOH_ID_num", breaks = c(360, 380, 400, 420, 440, 460, 480, 500), 
                     nsim = 1000)

# Print the power curve results
print(vin_pc)
plot(vin_pc)
```


## Analysis of the Letter and Dig Span

```{r analyzing the LDS}
# Run an empty model
m0_dls <- lmer(LetDig_Performance ~ (1 | HOH_ID), data = full_DLS)
performance::icc(m0_dls)
summary(m0_dls)

# Run the full model
m1_dls <- lmer(LetDig_Performance ~  Sex + Age_C + KBB_DD_status + (1 | HOH_ID), data = full_DLS)
summary(m1_dls)

# Run a model with just KBB_DD_Status
m2_dls <- lmer(LetDig_Performance ~  KBB_DD_status + (1 | HOH_ID), data = full_DLS)

# Calculate effect sizes (limitation)
t_to_eta2(0.874, 492.2633) # Inaccurate please address 

# Run a power analysis
powerSim(m2_dls)

# Extend the model to 500 clusters (adjust as needed)
cluster_n = 500
m3_dls <- extend(m2_dls, along = "HOH_ID_num", n = cluster_n) 

# Run powerCurve with explicit sample sizes using 'breaks'
dls_pc <- powerCurve(m3_dls, along = "HOH_ID_num", breaks = c(360, 380, 400, 420, 440, 460, 480, 500), 
                     nsim = 1000)

# Print the power curve results
print(dls_pc)
plot(dls_pc)
```


## Analysis of Pattern Reasoning

```{r analyzing the PR}
# Run an empty model
m0_pr <- lmer(PR_Performance ~ (1 | HOH_ID), data = full_PR)
performance::icc(m0_pr)
summary(m0_pr)

# Run the full model
m1_pr <- lmer(PR_Performance ~  Sex + Age_C + KBB_DD_status + (1 | HOH_ID), data = full_PR)
summary(m1_pr)

# Run a model with just KBB_DD_Status
m2_pr <- lmer(PR_Performance ~  KBB_DD_status + (1 | HOH_ID), data = full_PR)

# Calculate effect sizes (limitation)
t_to_eta2(-0.685, 433.6387) 

# Run a power analysis
powerSim(m2_pr)

# Extend the model to 500 clusters (adjust as needed)
cluster_n = 500
m3_pr <- extend(m2_pr, along = "HOH_ID_num", n = cluster_n) 

# Run powerCurve with explicit sample sizes using 'breaks'
pr_pc <- powerCurve(m3_pr, along = "HOH_ID_num", breaks = c(360, 380, 400, 420, 440, 460, 480, 500), 
                     nsim = 1000)

# Print the power curve results
print(pr_pc)
plot(pr_pc)
```

## Analysis of Triangles

```{r analyzing Triangles}
# Run an empty model
m0_tr <- lmer(TR_Performance ~ (1 | HOH_ID), data = full_TR)
performance::icc(m0_tr)
summary(m0_tr)

# Run the full model
m1_tr <- lmer(TR_Performance ~  Sex + Age_C + KBB_DD_status + (1 | HOH_ID), data = full_TR)
summary(m1_tr)

# Run a model with just KBB_DD_Status
m2_tr <- lmer(TR_Performance ~  KBB_DD_status + (1 | HOH_ID), data = full_TR)

# Calculate effect sizes (limitation)
t_to_eta2(1.471, 448.6302)  

# Run a power analysis
powerSim(m2_tr)

# Extend the model to 500 clusters (adjust as needed)
cluster_n = 500
m3_tr <- extend(m2_tr, along = "HOH_ID_num", n = cluster_n) 

# Run powerCurve with explicit sample sizes using 'breaks'
tr_pc <- powerCurve(m3_tr, along = "HOH_ID_num", breaks = c(360, 380, 400, 420, 440, 460, 480, 500), 
                     nsim = 1000)

# Print the power curve results
print(tr_pc)
plot(tr_pc)
```
## Analysis of BRIEF2 Parent

```{r analyzing BRIEF2}
# Run an empty model
m0_brp <- lmer(GEC ~ (1 | HOH_ID), data = full_BRIEF_P)
performance::icc(m0_brp)
summary(m0_brp)

# Run the full model
m1_brp <- lmer(GEC ~  Sex + Age_C + KBB_DD_status + (1 | HOH_ID), data = full_BRIEF_P)
summary(m1_brp)

# Run a model with just KBB_DD_Status
m2_brp <- lmer(GEC ~  KBB_DD_status + (1 | HOH_ID), data = full_BRIEF_P)

# Calculate effect sizes (limitation)
t_to_eta2(4.375, 441.4660) 

# Run a power analysis
powerSim(m2_brp)

# Extend the model to 500 clusters (adjust as needed)
cluster_n = 500
m3_brp <- extend(m2_brp, along = "HOH_ID_num", n = cluster_n) 

# Run powerCurve with explicit sample sizes using 'breaks'
m3_brp <- powerCurve(m3_brp, along = "HOH_ID_num", breaks = c(360, 380, 400, 420, 440, 460, 480, 500), 
                     nsim = 1000)

# Print the power curve results
print(m3_brp)
plot(m3_brp)
```

